---
title: "Assignment 9- Data 607"
author: "Saloua Daouki"
date: "2023-11-04"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction:

The goal of this asssignment is to choose one of [the New York Times APIs](https://developer.nytimes.com/apis), construct an interface in R to read in the JSON data, and transform it into an R DataFrame. 
So I picked two sections from the Top Stories API:

-books/review

-sunday review

## Loading Libraries:

The libraries will be used in this assignment are tidyverse for tidying the data, and jsonlite to import the data.

```{r }
library(tidyverse)
library(jsonlite)
library(stringr)
```

## Importing the data:
### Read the json files:

```{r pressure, echo=FALSE}
my_api_key <- "lC1J9EyT9ivGEdysMWXvgjd2D5bGQZMo"
url <- paste("https://api.nytimes.com/svc/topstories/v2/books/review.json?api-key=", my_api_key, sep='')
url1 <- paste("https://api.nytimes.com/svc/topstories/v2/sundayreview.json?api-key=",my_api_key, sep='' )
```

### Transform the json files into data frame:

```{r}
df_BooksReview <- as.data.frame(fromJSON(url))
head(df_BooksReview)
df_SundayReview <- as.data.frame(fromJSON(url1))
head(df_SundayReview)
```

### Interfacing the data:

```{r}
CombinedData <- rbind(df_BooksReview,df_SundayReview)

CombinedData <- CombinedData %>%
  select(results.title,results.abstract,results.url,results.created_date, results.published_date, last_updated,results.section,results.subsection, results.byline, results.des_facet)
head(CombinedData)
```
### Rename the columns of the combined data:

```{r}
colnames(CombinedData) = c("Title", "Abstract", "URL", "Created_Date", "Published_Date", "Last_Updated", "Section", "Subsection", "Author", "Filtered")
```

### Remove the string "By" from column Author:

```{r}
CombinedData$Author <- gsub("By", "", as.character((CombinedData$Author)))
```

### Breaking the column "Filtered" into a list:

```{r}
DF_Reviews <- unnest(CombinedData, Filtered)
head(DF_Reviews)
```

### Counting the frequency of each facet in "Filtered" column:

```{r}
freq_DF_Reviews <- as.data.frame(table(DF_Reviews$Filtered))
colnames(freq_DF_Reviews)[1]="Facets"
glimpse(freq_DF_Reviews)
```






